Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-07-01 paper/paper.tex
update the figure

       \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} & & & &
       \begin{smallmatrix} \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
     };
   {
     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-4);
     \chainin (m-2-5) [join={node[sloped,anchor=center,above,labeled] {\text{CPS left-inv.}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-5);
     \chainin (m-2-4) [join={node[sloped,anchor=center,below,labeled] {\text{CPS trans.}}}];
   }
   {
     [start chain, color=gray] \chainin (m-1-5);
     \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
     \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
     \chainin (m-2-4) [join={node[sloped,anchor=center,above,labeled] {\text{inline monads}}}];
   }
   {
     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-1);
     \chainin (m-2-2) [join={node[above,labeled] {
       \begin{smallmatrix} \text{linearization} \\ \text{+} \\ \text{lightw. fusion} \end{smallmatrix}
     }}];
     \chainin (m-2-3) [join={node[above,labeled] {\text{disentang.}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-2);
     \chainin (m-2-1) [join={node[below,labeled] {\begin{smallmatrix} \text{splitting} \\ \text{+} \\ \text{List(Success)??} \end{smallmatrix}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-3);
     \chainin (m-2-2) [join={node[below,labeled] {\text{entang.}}}];
   }
   {
     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-3);
     \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-4);
     \chainin (m-2-3) [join={node[below,labeled] {\text{defunc.}}}];
   }
   {
     [start chain, color=gray, dashed] \chainin (m-3-1);
     { [start branch=A] \chainin (m-2-1) [join={node[left, labeled] {\alpha}}]; }
   }
   {
     [start chain, color=gray, dashed] \chainin (m-3-5);
     { 
       [start branch=B] \chainin (m-2-5)
       [join={node[right, labeled] {\alpha}}];
     }
     {
       [start branch=D, color=gray] \chainin (m-3-1)
       [join={node[right, labeled] {\text{}}}];
     }
   }
   {
     [start chain, color=gray] \chainin (m-3-1);
     {
       [start branch=C] \chainin (m-3-5)
       [join={node[above, labeled] {\text{Known as Danvy's correspondences for concrete semantic artifacts }}}];
     }
   }
   \draw[gray,dashed,transform canvas={xshift=0.5ex}] (m-1-5.north west) -- (m-2-5.south west);
   \draw[gray,dashed,transform canvas={xshift=3.7ex, xscale=0.90}] (m-1-5.north west) -- (m-1-5.north east);
   \draw[gray,dashed,transform canvas={xshift=-0.2ex}] (m-1-5.north east) -- (m-2-5.south east);
   \draw[gray,dashed,transform canvas={xshift=3.7ex, xscale=0.90}] (m-2-5.south west) -- (m-2-5.south east);
 \end{tikzpicture}
 \caption{Functional correspondence between and transformations from AAM to ADI} \label{fig:trans}
\end{center}
\end{figure}

@@

At this point, we have obtained a linearized abstract abstract machine. 
Intuitively, if we imagine that the classical AAM explores a graph of reachable states, 
then the linearized AAM flattens the graph to a linear sequence of states.
It has been observed that ``\textit{failure can be replaced by a list of successes}''
\cite{10.1007/3-540-15975-4_33} and that ``\textit{one programmer’s lazy-list 
constructor is another programmer’s success and failure continuations}'' \cite{danvy2006a}; 
the @step@ function in AAM is the list constructor used in a lazy way, and here we turn it 
into an explicit data types representing continuation.
The inverse transformation of this step is to apply \citeauthor{10.1007/3-540-15975-4_33}'s 
technique that converts the data types representing meta-continuations (failure continuations) 
back to list representation, where the elements in the list are plain machine states.

>>>>

       \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} & & & &
       \begin{smallmatrix} \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
     };
   {
     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-4);
     \chainin (m-2-5) [join={node[sloped,anchor=center,above,labeled] {\text{CPS left-inv.}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-5);
     \chainin (m-2-4) [join={node[sloped,anchor=center,below,labeled] {\text{CPS trans.}}}];
   }
   {
     [start chain, color=gray] \chainin (m-1-5);
     \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
     \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
     \chainin (m-2-4) [join={node[sloped,anchor=center,above,labeled] {\text{inline monads}}}];
   }
   {
     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-1);
     \chainin (m-2-2) [join={node[above,labeled] {
       \begin{smallmatrix} \text{linearization} \\ \text{+} \\ \text{lightw. fusion} \end{smallmatrix}
     }}];
     \chainin (m-2-3) [join={node[above,labeled] {\text{disentang.}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-2);
     \chainin (m-2-1) [join={node[below,labeled] {\begin{smallmatrix} \text{splitting} \\ \text{+} \\ 
     \text{replacing cont.} \\ \text{by list} \end{smallmatrix}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-3);
     \chainin (m-2-2) [join={node[below,labeled] {\text{entang.}}}];
   }
   {
     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-3);
     \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
   }
   {
     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-4);
     \chainin (m-2-3) [join={node[below,labeled] {\text{defunc.}}}];
   }
   {
     [start chain, color=gray, dashed] \chainin (m-3-1);
     { [start branch=A] \chainin (m-2-1) [join={node[left, labeled] {\alpha}}]; }
   }
   {
     [start chain, color=gray, dashed] \chainin (m-3-5);
     { 
       [start branch=B] \chainin (m-2-5)
       [join={node[right, labeled] {\alpha}}];
     }
     {
       [start branch=D, color=gray] \chainin (m-3-1)
       [join={node[right, labeled] {\text{}}}];
     }
   }
   {
     [start chain, color=gray] \chainin (m-3-1);
     {
       [start branch=C] \chainin (m-3-5)
       [join={node[above, labeled] {\text{Known as Danvy's correspondences for concrete semantic artifacts }}}];
     }
   }
   \draw[gray,dashed,transform canvas={xshift=0.5ex}] (m-1-5.north west) -- (m-2-5.south west);
   \draw[gray,dashed,transform canvas={xshift=3.7ex, xscale=0.90}] (m-1-5.north west) -- (m-1-5.north east);
   \draw[gray,dashed,transform canvas={xshift=-0.2ex}] (m-1-5.north east) -- (m-2-5.south east);
   \draw[gray,dashed,transform canvas={xshift=3.7ex, xscale=0.90}] (m-2-5.south west) -- (m-2-5.south east);
 \end{tikzpicture}
 \caption{Functional correspondence between and transformations from AAM to ADI} \label{fig:trans}
\end{center}
\end{figure}

@@

At this point, we have obtained a linearized abstract abstract machine. 
Intuitively, if we imagine that the classical AAM explores a graph of reachable states, 
then the linearized AAM flattens the graph to a linear sequence of states.
It has been observed that ``\textit{failure can be replaced by a list of successes}''
\cite{10.1007/3-540-15975-4_33} and that ``\textit{one programmer’s lazy-list 
constructor is another programmer’s success and failure continuations}'' \cite{danvy2006a}; 
the @step@ function in AAM is the list constructor used in a lazy way, and here we turn it 
into an explicit data types representing continuation.
The inverse transformation of this step is to apply \citeauthor{10.1007/3-540-15975-4_33}'s 
technique that converts the data types representing meta-continuations (failure continuations) 
back to lazy list representation, where the elements in the list are plain machine states.
